---
title: "P8105 Homework 1"
author: "Alina Levine"
output: github_document
---

# Problem 1


I am creating a random sample of size 10 from the uniform distribution and will make 
a logical vector indicating which elements are greater than 2. I will also create a factor vector
and character vector

```{r create 4 types of vectors}

uniform_sample_vector = runif(10, 0, 5)

greater_than_two = uniform_sample_vector > 2

factor_vector = as.factor(c("a", "a", "a", "a", "a", "a", "b", "b", "b", "b"))

character_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j")

```

When I take the mean of these vectors, I get a number back for the 
logical and numeric vctors, but I get NA for the factor and character vectors. This is because
mean() treats the logical vector as numeric, giving TRUE a value of 1
and FALSE a value of 0, but it does not do this with character vectors

```{r take means}
mean(uniform_sample_vector)
mean(greater_than_two)
mean(factor_vector)
mean(character_vector)

```


When I apply as.numeric to the logical vector, I get a vector of 1s and 0s return
When I apply as.numeric to the factor vector, I get a vector of 1 and 2, each number
corresponding to a level. I get NAs when I apply as.numeric to a character vector

```{r convert to numeric, results = hide}

as.numeric(greater_than_two)
as.numeric(factor_vector)
as.numeric(character_vector)


```

When I convert my character vector to a factor and then to numeric I get a numeric vector 1:10
because the character vector had 10 unique entries, so there were 10 levels. When I convert my 
factor vector to a character vector to a numeric vector I get NAs because character vectors 
cannot be converted to numeric vectors

```{r converting between three vector types, results = hide}

as.numeric(as.factor(character_vector))
as.numeric(as.character(factor_vector))


```




