---
title: "P8105 Homework 1"
author: "Alina Levine"
output: github_document
---

```{r packages}
library(tidyverse)
library(ggplot2)
```



# Problem 1

## Create Vectors
I am creating a random sample of size 10 from the uniform distribution and will make 
a logical vector indicating which elements are greater than 2. I will also create a factor vector
and character vector

```{r create 4 types of vectors}

uniform_sample_vector = runif(10, 0, 5)

greater_than_two = uniform_sample_vector > 2

factor_vector = as.factor(c("a", "a", "a", "a", "a", "a", "b", "b", "b", "b"))

character_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j")

```
##Mean of Vectors
When I take the mean of these vectors, I get a number back for the 
logical and numeric vctors, but I get NA for the factor and character vectors. This is because
mean() treats the logical vector as numeric, giving TRUE a value of 1
and FALSE a value of 0, but it does not do this with character vectors

```{r take means, results = "hide", message = FALSE}
mean(uniform_sample_vector)
mean(greater_than_two)
mean(factor_vector)
mean(character_vector)

```

## Converting to Numeric
When I apply as.numeric to the logical vector, I get a vector of 1s and 0s return
When I apply as.numeric to the factor vector, I get a vector of 1 and 2, each number
corresponding to a level. I get NAs when I apply as.numeric to a character vector

```{r convert to numeric, results= "hide"}

as.numeric(greater_than_two)
as.numeric(factor_vector)
as.numeric(character_vector)


```

When I convert my character vector to a factor and then to numeric I get a numeric vector 1:10
because the character vector had 10 unique entries, so there were 10 levels. When I convert my 
factor vector to a character vector to a numeric vector I get NAs because character vectors 
cannot be converted to numeric vectors

```{r converting between three vector types, results = "hide"}

as.numeric(as.factor(character_vector))
as.numeric(as.character(factor_vector))


```

#Problem 2


##Create Data Frame
```{r, create data frame}

set.seed(1)

plot_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  x_plus_y_larger_zero = (x + y) > 0,
  x_plus_y_larger_zero_numeric = as.numeric(x_plus_y_larger_zero),
  x_plus_y_larger_zero_factor = as.factor(x_plus_y_larger_zero)
)

```



There are `r ncol(plot_df)` columns and `r nrow(plot_df)` rows.  
The mean of x is `r mean(plot_df$x)` and the median of x is `r median(plot_df$x)`  
The proportion of cases for which x_plus_y_larger_zero is TRUE is `r sum(plot_df$x_plus_y_larger_zero)/length(plot_df$x_plus_y_larger_zero)`




##Scatter Plots

I get the same result when I color by a factor and by a logical vector. When I color by a numerical vector the only difference I see is that the legend shows a gradient of color going from 0 to 1 because the numbers in the vectors range from 0 to 1 since FALSE gets 0 and TRUE gets 1

```{r}
ggplot(plot_df, aes(x = x, y = y, color = x_plus_y_larger_zero)) + labs(title = "Colored by Logical Vector") + geom_point()


ggplot(plot_df, aes(x = x, y = y, color = x_plus_y_larger_zero_numeric )) + labs(title = "Colored by Numeric Vector") + geom_point()


ggplot(plot_df, aes(x = x, y = y, color = x_plus_y_larger_zero_factor )) + labs(title = "Colored by Factor Vector") + geom_point()





```










